%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%A  options.tex    ANUPQ documentation - options                Werner Nickel
%A                                                                Greg Gamble
%%
%A  @(#)$Id$
%%
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{ANUPQ Options}

In this chapter we describe in detail all the options used  by  functions
of the {\ANUPQ} share package. Note that by ``options''  we  mean  {\GAP}
options that are passed to functions after the  arguments  and  separated
from the arguments by  a  colon  as  described  in  Chapter~"ref:Function
Calls" in the Reference Manual. The user  is  strongly  advised  to  read
Section~"Hints and Warnings regarding the use of Options".

\>AllANUPQoptions() F

lists all the {\GAP} options defined for functions of the {\ANUPQ}  share
package:

\beginexample
gap> AllANUPQoptions();
[ "AllDescendants", "BasicAlgorithm", "Bounds", "CapableDescendants", 
  "ClassBound", "Exponent", "GroupName", "Metabelian", "OrderBound", 
  "OutputLevel", "PcgsAutomorphisms", "PqWorkspace", "Prime", "QueueFactor", 
  "RankInitialSegmentSubgroups", "Relators", "SetupFile", "SpaceEfficient", 
  "StandardPresentationFile", "StepSize", "SubList", "TailorOutput" ]
\endexample

The following global variable gives a  partial  breakdown  of  where  the
above options are used.

\>`ANUPQoptions' V

is a record of lists of names of admissible {\ANUPQ} options, such  that
each field is either the name of a ``key'' {\ANUPQ} function  or  `other'
(for a miscellaneous list of functions) and the  corresponding  value  is
the list of option  names  that  are  admissible  for  the  function  (or
miscellaneous list of functions).

Also, from within a {\GAP} session, you may  use  {\GAP}'s  help  browser
(see Chapter~"ref:The Help System" in the {\GAP}  Reference  Manual);  to
find out about any particular {\ANUPQ} option,  simply  type:  ```?option
<option>''', where <option> is one of the options  listed  above  without
any quotes, e.g.

\begintt
gap> ?option Prime
\endtt

will display the section  in  this  manual  that  describes  the  `Prime'
option.

In the section following we describe in detail all {\ANUPQ}  options.  To
continue onto the next section on-line using {\GAP}'s help browser, type:

\begintt
gap> ?>
\endtt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Detailed descriptions of ANUPQ Options}

\beginitems

\>`Prime := <p>'{option Prime}@{option `Prime'|indexit}&
Specifies that the $p$-quotient for the prime  <p>  should  be  computed.

\>`ClassBound := <n>'{option ClassBound}@{option `ClassBound'|indexit}&
Specifies that the $p$-quotient computed has lower exponent-$p$ class  at
most <n>.

\>`Exponent := <n>'{option Exponent}@{option `Exponent'|indexit}&
Specifies that the $p$-quotient computed has exponent <n>. By default, no
exponent law is enforced.

\>`Relators := <rels>'{option Relators}@{option `Relators'|indexit}&
Specifies that the relators sent to the  `pq'  binary  should  be  <rels>
rather than the relators of the group <F>; <rels> should  be  a  list  of
*strings* in the string representations of the generators  of  <F>.  This
option provides a way of giving relators  to  the  `pq'  binary,  without
having  them  pre-expanded  by  {\GAP},  which  can  sometimes  effect  a
performance loss of the order of 100.

*Note:*
The `pq' binary does not use `/' to indicate multiplication by an inverse
and uses square brackets to represent (left norm) commutators. Also, even
though the `pq' binary accepts relations, all elements of  <rels>  *must*
be in relator form, i.e.~a relation of form `<w1> = <w2>' must be written
as `<w1>*(<w2>)^-1' and then put in a pair of double-quotes to make it  a
string. See the example below.

\>`Metabelian'{option Metabelian}@{option `Metabelian'|indexit}&
Specifies that the largest metabelian  $p$-quotient  be  constructed.  By
default this restriction is not enforced.

\>`GroupName := <name>'{option GroupName}@{option `GroupName'|indexit}&
Specifies that the `pq' binary should refer to  the  group  by  the  name
<name> (a string). If `GroupName' is not  set  and  the  group  has  been
assigned a name via `SetName' (see~"ref:SetName") it is set as  the  name
the `pq' binary should use. Otherwise, the ``generic'' name `"\<grp>"' is
set as a default.

\>`OutputLevel := <n>'{option OutputLevel}@{option `OutputLevel'|indexit}&
Specifies the level of ``verbosity'' of the information output by the ANU
`pq' binary when computing a pc presentation; <n> must be an  integer  in
the range 0 to 3. `OutputLevel := 0' displays at most one line of  output
and is the default; `OutputLevel := 1' displays (usually)  slightly  more
output and `OutputLevel's of 2 and 3 are two levels of verbose output. To
see these messages from the `pq' binary, the `InfoANUPQ'  level  must  be
set to at least 1 (see~"InfoANUPQ").

\>`SetupFile := <filename>'{option SetupFile}@{option `SetupFile'|indexit}&
Non-interactively, this option directs that `pq' should not be called and
that an input file  with  name  <filename>  (a  string),  containing  the
commands necessary for the  ANU  `pq'  standalone,  be  constructed.  The
commands written to <filename> are also `Info'-ed behind  a  ```ToPQ> '''
prompt at `InfoANUPQ' level  4  (see~"InfoANUPQ").  In  this  case,  `Pq'
returns `true'. Interactively, `SetupFile' is ignored.

\>`PqWorkspace := <workspace>'{option PqWorkspace}@{option `PqWorkspace'|indexit}&
Non-interactively, this option sets the memory used by the  `pq'  binary.
It sets the maximum number of integer-sized elements to allocate  in  its
main storage array. By default, the  `pq'  binary  sets  this  figure  to
10000000. Interactively, `PqWorkspace' is ignored;  the  memory  used  in
this  case  may  be  set  by   giving   `PqStart'   a   second   argument
(see~"PqStart").

\>`PcgsAutomorphisms'{option PcgsAutomorphisms}@{option `PcgsAutomorphisms'|indexit}&
Specifies that a polycyclic  generating  sequence  for  the  automorphism
group of <G> (which must be *soluble*), be computed  and  passed  to  the
`pq' binary. This increases the efficiency of the  computation;  it  also
prevents the `pq' from calling {\GAP} for orbit-stabilizer  calculations.
See  Section~"Computing   Descendants   of   a   p-Group"   for   further
explanations.

*Note:*
If `PcgsAutomorphisms' is used when the  automorphism  group  of  <G>  is
insoluble, an error message occurs.

\>`OrderBound := <n>'{option OrderBound}@{option `OrderBound'|indexit}&
Specifies that only descendants of size at most $p^<n>$, where <n>  is  a
non-negative integer,  be  generated.  Note  that  you  cannot  set  both
`OrderBound' and `StepSize'.

\>`StepSize := <n>'{option StepSize}@{option `StepSize'}
\>`StepSize := <list>'{option StepSize}@{option `StepSize'|indexit}&
For  a  positive  integer  <n>,  `StepSize'  specifies  that  only  those
immediate descendants which are a factor $p^<n>$ bigger than their parent
group be generated.

&
For a list <list> of positive integers such that the sum of the length of
<list> and the exponent-$p$ class of <G> is  equal  to  the  class  bound
defined  by  the  option  `ClassBound',  `StepSize'  specifies  that  the
integers of <list> are the step sizes for each additional class.

\>`RankInitialSegmentSubgroups := <n>'{option RankInitialSegmentSubgroups}%
@{option `RankInitialSegmentSubgroups'|indexit}&
Sets the rank of the initial  segment  subgroup  chosen  to  be  <n>.  By
default, this has value 0.

\>`SpaceEfficient'{option SpaceEfficient}@{option `SpaceEfficient'|indexit}&
Specifies that the `pq' binary perform calculations more slowly but  with
greater space efficiency. This flag is frequently necessary for groups of
large Frattini quotient rank. The space saving occurs  because  only  one
permutation is stored at any one time. This option is only  available  in
conjunction with the `PcgsAutomorphisms' flag.

\>`CapableDescendants'{option CapableDescendants}@{option `CapableDescendants'|indexit}&
By default, *all* (i.e.~capable and terminal) descendants  are  computed.
If this flag is set, only capable descendants are computed. Setting  this
option is equivalent to setting `AllDescendants  :=  false'  (see~"option
AllDescendants"),    except    if    both    `CapableDescendants'     and
`AllDescendants' are passed, `CapableDescendants' takes precedence.

\>`AllDescendants := false'{option AllDescendants}@{option `AllDescendants'|indexit}&
By default, *all* descendants are constructed. If this  flag  is  set  to
`false', only capable descendants are computed. Passing  `AllDescendants'
without a value  (which  is  equivalent  to  setting  it  to  `true')  is
superfluous. This option is provided only for backward compatibility with
the {\GAP} 3 version of the {\ANUPQ}  share  package,  where  by  default
`AllDescendants'  was  set  to  `false'  (rather  than  `true').  It   is
preferable to use `CapableDescendants' (see~"option CapableDescendants").

\>`SubList := <sub>'{option SubList}@{option `SubList'|indexit}&
Suppose that <L> is the list of descendants generated, then  for  a  list
<sub> of integers this option causes `PqDescendants' to return  `Sublist(
<L>, <sub> )'. If an integer <n>  is  supplied,  `PqDescendants'  returns
`<L>[<n>]'.

\>`BasicAlgorithm'{option BasicAlgorithm}@{option `BasicAlgorithm'|indexit}&
Specifies that an algorithm that the `pq' binary  calls  its  ``default''
algorithm be used. By default this  algorithm  is  *not*  used.  If  this
option is supplied the settings of options `RankInitialSegmentSubgroups',
`AllDescendants', `Exponent' and `Metabelian' are ignored.

\>`TailorOutput := <rec>'{option TailorOutput}@{option `TailorOutput'|indexit}&
Specifies that fine tuning of the output is  desired.  The  record  <rec>
should have any subset (or all) of the the following fields:

\quad`perm := <list>' &
where <list> is a list of booleans which determine (in order) whether one
wants to print for the permutation group output: the degree, the extended
automorphisms, the automorphism matrices, and the permutations.

\quad`orbit := <list>' &
where <list> is a list of booleans which determine (in order) whether one
wants to print for the orbit output: a summary, and a complete listing of
orbits. (It's possible to have *both* a summary and a complete listing.)

\quad`group := <list>' &
where <list> is a list of booleans which determine (in order) whether one
wants to print  for  the  group  output:  the  standard  matrix  of  each
allowable subgroup, the presentation of reduced $p$-covering groups,  the
presentation of immediate descendants, the nuclear rank  of  descendants,
and the $p$-multiplicator rank of descendants.

\quad`autgroup := <list>' &
where <list> is a list of booleans which determine (in order) whether one
wants to print for the automorphism group output: the commutator  matrix,
the automorphism group description of descendants, and  the  automorphism
group order of descendants.

\quad`trace := <val>' &
where <val> is a boolean which if `true' specifies algorithm  trace  data
is desired. By default, one does not get algorithm trace data.

Not providing a field (or mis-spelling it!), specifies that  the  default
output is desired. As a convenience, `1' is also accepted as `true',  and
any value that is neither `1' nor `true' is taken as  `false'.  Also  for
each <list> above, an unbound list entry is taken as `false'.  Thus,  for
example

\begintt
TailorOutput := rec(group := [,,1], autgroup := [,1])
\endtt

specifies for the group output that only the  presentation  of  immediate
descendants is desired,  for  the  automorphism  group  output  only  the
automorphism group description of descendants  should  be  printed,  that
there should be no algorithm trace data,  and  that  the  default  output
should be provided for the permutation group and orbit output.

\>`StandardPresentationFile := <filename>'{option StandardPresentationFile}%
@{option `StandardPresentationFile'|indexit}&
Specifies that the file to which the standard presentation is written has
name <filename>. If the first character of the string <filename>  is  not
`/', <filename> is assumed to be the path of a writable file relative  to
the directory in which {\GAP} was started. If this option is  omitted  it
is written to the file with the name generated by the command  `Filename(
ANUPQData.tmpdir, "SPres" );', i.e.~the file with name `"SPres"'  in  the
temporary directory in which the `pq' binary executes.

\>`QueueFactor := <n>'{option QueueFactor}@{option `QueueFactor'|indexit}&
Specifies a queue factor of <n>, where <n> should be a positive  integer.
This option may be used with `PqNextClass' (see~"PqNextClass").

\>`Bounds := <list>'{option Bounds}@{option `Bounds'|indexit}&
Specifies a lower and upper bound on the indices of a list, where  <list>
is a pair of positive non-decreasing  integers.  See~"PqDisplayStructure"
and~"PqDisplayAutomorphisms" where this option may be used.

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%E
